---
title: "Riddler - Skip to next song"
author: Ravi Chandrasekaran
date: December 9, 2019
output: html_document
---

https://fivethirtyeight.com/features/how-fast-can-you-skip-to-your-favorite-song/
https://en.wikipedia.org/wiki/Negative_binomial_distribution#Waiting_time_in_a_Bernoulli_process

```{r package_setup, echo = FALSE}
options(tidyverse.quiet = TRUE) 
library("tidyverse")
```


## General approach
When we are very close to our desired target, our gut tells us we should
be patient and hit "next" a few times. When we're ridiculously far away, our
same gut says to hit "random" a few times, and hope that we get close. The
question is how close is "close" -- when should we be patient, and when 
should we try for better.

## Closeness
Let's say our threshold for closeness is $n$ -- that we are either on our
desired square, or in the other $n - 1$ nearest, lesser squares. For example,
if our target is 42, and our threshold was $n = 10$, then the "close" squares
would include $33, 34, \dots, 42$. 

If we land in the close squares, with all probabilities equal, then how many
times would we expect to hit the next button? Well, if it's an equal chance
for each of these close squares, then the expectation is a simple calc:

\begin{align*}
E[F] = & \frac{1}{2}(n - 1) \\
\end{align*}


## Random steps
How many times would we expect to hit the "random" button before landing in the
"close" zone? 

Let's start by assuming the starting position is outside the "close" zone. 
We're basically looking at a geometric distribution, then, where the 
probability of failure and success is driven by the size of the closeness
threshold. 

What's our chance ($p$) -- of landing in the close zone? That's
defined by our closeness threshold and the total number of squares we're 
considering -- $p = n / N$. So, we can build up an equation entirely based 
on the cutoff and total number of squares, below:

\begin{align*}
p = & \frac{n}{N} \\
E[R \mid \text{not close}] = & \frac{1}{p}, && \text{from the geometric dist.} \\
= & \frac{N}{n} \\
\end{align*}

Now, we had started by assuming we were *outside* the close zone. Of course, 
if our starting position were lucky enough, we wouldn't have any random steps
at all. So:

\begin{align*}
E[R] = & (1 - p) \cdot E[R \mid \text{not close}] + p \cdot E[R \mid \text{close}] \\
= & (1 - \frac{n}{N}) \cdot \frac{N}{n} + \frac{N}{n} \cdot 0 \\
= & (\frac{N - n}{N}) \cdot \frac{N}{n}  \\
= & \frac{N - n}{n} \\
= & \frac{N}{n} - 1 \\
\end{align*}

## Putting it together
Our total number of expected steps is the sum of the expected random button 
presses, and the expected number of forward button presses. These are both
functions of the closeness threshold $n$, so once we have the formula, we'll
want to find the $n$ that minimizes the number of steps.

\begin{align*}
E[S] = & E[R] + E[F] \\
= & \frac{N}{n}  - 1 + \frac{1}{2}(n - 1) \\
= & \frac{N}{n} +  \frac{n}{2}  + \frac{1}{2} \\
\end{align*}

This is enough for us to plot the expected steps at each value of the closeness
threshold $n$. Here, we're first plotting all values, and then limiting it to 
just those values close to the minimum. From here, we can eyeball the minimum
at $n = 14$.


```{r plot_exp_steps, echo = FALSE}
library(ggrepel)
exp_random_steps = function(n) { ((100 - n) / n) }
exp_forward_steps = function(n) { 1/2 * (n - 1) }


df <- tibble(
  threshold = seq(1, 100),
  random_steps = exp_random_steps(seq(1, 100)),
  forward_steps = exp_forward_steps(seq(1, 100))
  ) %>%
  mutate(steps = random_steps + forward_steps)

df %>%
  filter(between(threshold, 10, 20)) %>%
  ggplot(aes(threshold, steps)) +
  geom_point(aes(col = steps == min(steps)), show.legend = FALSE) +
  labs(title = "Expected steps at each 'closeness' threshold",
       x = "'Closeness' threshold",
       y = "Expected steps") +
  theme_light() +
  scale_x_continuous(
    breaks = seq(0, 100), 
    minor_breaks = seq(0, 100)) +
  scale_color_manual(
    labels = c("Minimum", "Not minimum"), 
    values = c("purple", "orange"),
    guide = guide_legend(reverse = TRUE)) +
  geom_text_repel(
    data = subset(df, steps == min(steps)), 
    aes(label = round(steps, 1)),
    size = 3) +
  coord_cartesian(ylim = c(12, 14))
```


At our minimizing threshold, we get a pretty balanced mix of expected 
random and forward steps:

```{r breakdown_at_ideal_cutoff, echo = FALSE}

df %>%
  filter(threshold == 14) %>%
  select(-threshold) %>%
  gather(key = "metric") %>%
  mutate(metric = str_to_sentence(str_replace_all(metric, "_", " ")))

```

Here we can see how that balance plays out for each possible value of the
closeness threshold $n$. Not unexpectedly, when we give a crazy large threshold
for closeness, the "next" buttons dominate, and that's a linear trend.

```{r plot_balance, echo = FALSE}
df %>%
  select(-steps) %>%
  gather(key = "type", value = "steps", -threshold) %>%
  ggplot(aes(threshold, steps, fill = type)) +
  geom_col() + 
  theme_light() +
  theme(
    legend.position = c(.85, .85),
    legend.title = element_blank()) +
  labs(title = "Expected steps at each 'closeness' threshold",
       x = "'Closeness' threshold",
       y = "Expected steps") +
  scale_x_continuous(breaks = seq(0, 100, 20)) +
  scale_y_continuous(breaks = seq(0, 100, 20)) +
  scale_fill_manual(
    labels = c("Forward", "Random"), 
    values = c("grey80", "lightblue"),
    guide = guide_legend(reverse = TRUE))
  
```



## Rule of thumb

What do we do if our number of songs isn't $N = 100$, but some other value?
We can use a little calc on our expected steps function to differentiate and 
find the minimum of expected steps as a function of that value $N$.

\begin{align*}
E[S] = & \frac{N}{n} - 1 + \frac{1}{2}(n - 1) \\

\frac{d}{dn}E[S] = & \frac{-N}{n^2} + 0 + \frac{1}{2} \\
0 = & \frac{-N}{n^2} + \frac{1}{2} \\
n^2 = & 2N \\
n = & \sqrt{2N} \\
\end{align*}

\begin{align*}
= & \sqrt{2 \cdot 100} \\
= & `r round(sqrt(2 * 100), 2)` \\
\end{align*}


---


## Does this hold up under trials?
Enough maths. Let's try doing a whole bunch of trials with different
cutoffs, and see what we find. 

```{r perform_simulation, echo = FALSE, eval = FALSE}
# Currently ignoring boundary condition, since threshold and target don't 
# require it.
steps <- function(x, threshold, target) {
  if (x == target) { return(0) }
  if (target - x >= 0 & target - x < threshold) { return(target - x) }
  return(steps(as.integer(runif(1, min = 1, max = 100)), threshold, target) + 1)
}

trials <- expand_grid(
  start = sample(1:100, size = 10000, replace = TRUE),
  threshold = 5:30, 
  target = 42) 

trial_detail <- trials %>%
  mutate(row_number = row_number()) %>%
  nest(data = c(threshold, target, start)) %>%
  mutate(steps = map(data, ~ steps(.x$start, .x$threshold, .x$target))) %>%
  unnest(c(data, steps))
```

```{r load_simulation, eval = TRUE, echo = FALSE}
# save(trial_detail, file = "riddler-2019-12-06-trial-detail.Rds")
load("riddler-2019-12-06-trial-detail.Rds")
```


```{r evaluate_simulation, echo = FALSE}
trial_summary <- trial_detail %>%
  group_by(threshold) %>%
  summarise(steps = mean(steps))

trial_summary %>%
  filter(between(threshold, 10, 20)) %>%
  ggplot(aes(threshold, steps)) +
  geom_point(aes(col = steps == min(steps)), show.legend = FALSE) +
  labs(title = "Experimental steps for each closeness threshold",
       x = "'Closeness' threshold",
       y = "Mean number of steps") +
  theme_light() +
  coord_cartesian(ylim = c(12, 14)) +
  scale_x_continuous(
    breaks = seq(0, 100), 
    minor_breaks = seq(0, 100)) +
  labs(title = "Expected steps at each 'closeness' threshold",
       x = "'Closeness' threshold",
       y = "Expected steps") +
  scale_color_manual(
    labels = c("Minimum", "Not minimum"), 
    values = c("purple", "orange"),
    guide = guide_legend(reverse = TRUE)) +
  geom_text(
    data = subset(trial_summary, abs(steps - min(steps)) <= .1), 
    aes(label = round(steps, 1)),
    nudge_y = .1,
    size = 3) 
  

```

---



```{r ridgeplot, echo = FALSE}
library("ggridges")

trial_detail %>%
  filter(between(threshold, 10, 20)) %>%
  ggplot(aes(steps, y = threshold, group = threshold)) +
  geom_density_ridges(scale = 3, fill = "lightblue", color = "grey50") +
  theme_light() +
  labs(
    title = "Densities for number of steps",
    x = "Number of steps",
    y = "Density") +
  scale_y_continuous(breaks = seq(10, 20, 2)) +
  coord_cartesian(xlim = c(0, 50))
```


```{r variance_trials, echo = FALSE}
library("broom")

# Later is lower variance
trial_detail %>%
  select(-row_number, -target, -start) %>%
  nest(data = c(steps)) %>%
  mutate(t_test = purrr::map(data, ~ broom::tidy(t.test(.x$steps)))) %>%
  unnest(t_test) %>%
  ggplot(aes(threshold, estimate)) + 
  geom_point() +
  geom_linerange(aes(ymin = conf.low, ymax = conf.high)) +
  theme_light() +
  labs(
    title = "Point estimates and confidence intervals for expected steps", 
    x = "'Closeness' threshold",
    y = "Number of steps")

```










